---
title: "Analysis of BSL corpus questions"
author: "Bodo Winter"
date: "02/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Analysis of questions in the BSL corpus. Sections labeled **turning point** exclude data which means that everything from that section onward will be a subset analysis.

The analysis has two turning points:
- `semantic types`: from that point onwards, all analyses focus on polar versus content questions
- `constructed actions`: from that point onwards, all analyses focus on the non-CA dataset

Load packages:

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(lsr) # for Cramer's V effect size
library(patchwork) # for multiplotting
```

Check R version and package version for reporting:

```{r}
R.Version()
packageVersion('tidyverse')
packageVersion('lsr')
packageVersion('patchwork')
```

Load helper functions. Loading this makes the new functions `proportion()` and `percentage()` available, which I wrote because this project requires lots of calculation of proportions and percentages on the basis of contigency tables. The custom function `prop_table()` then uses both of these functions to create a table of proportions and percentages. It takes an object resulting from `count()` as input.

Sourcing this .R script also makes the ggplot2 theme `theme_BSL()` available, which is an extension of `theme_classic()` with some project-specific cosmetic tweaks.

```{r}
source('question_helper_functions.R')
```

Load cleaned data, which is the output from the `preprocessing.Rmd` script:

```{r, message = FALSE, warning = FALSE}
df <- read_csv('../data/BSL_questions_cleaned.csv')
```

## CA by form

Check whether the constructed actions stand out from the non-CA forms in any of the major formal characteristics. For this we'll use standardized residuals:

Manual:

```{r}
xtab <- with(df, table(has_CA, manual))

# Check:

xtab
prop_table(xtab)
stdres(xtab)
```

Forms with constructed actions more likely to contain single manual signs (+4.8), and less likely to contain none (-4.9).

Non-manual:

```{r}
xtab <- with(df, table(has_CA, nonmanual))

# Check:

xtab
prop_table(xtab)
stdres(xtab)
```

No difference with respect to nonmanuals.

```{r}
xtab <- with(df, table(has_CA, mouthing_question))

# Check:

xtab
prop_table(xtab)
stdres(xtab)
```

Constructed actions much more likely (+4.8) to have single mouthings, and much less likely to have no mouthing (-4.6).

```{r}
xtab <- with(df, table(has_CA, mouth_gesture_question))

# Check:

xtab
prop_table(xtab)
stdres(xtab)
```

Constructed actions more likely to have mouth gestures (+2.2) than not (-2.2).

For eye brow, use the simplified version (up/down/neither):

```{r}
eye_simple_df <- df %>%
  filter(eye_n_brow %in% c('up', 'down', 'null')) %>% 
  mutate(eye_n_brow = str_replace(eye_n_brow,
                                  'null', 'neither'))

# Check:

xtab <- with(eye_simple_df, table(has_CA, eye_n_brow))

# Check:

xtab
prop_table(xtab)
stdres(xtab)
```

More likely "down" (+2.3) for CA, but no particularly strong patterns for "neither" and "up" to contrast with.

Same for head movements (using the simplified/cleaned variable to avoid having too many categories).

```{r}
head_df <- df %>% 
  mutate(head = ifelse(head == 'Nod1',
                       'nod', head)) %>% 
  filter(head %in% c('Null', 'Forward', 'Nod',
                     'Nod1', 'Sideward', 'Down'))

# Check:

xtab <- with(head_df, table(has_CA, head))

# Check:

xtab
prop_table(xtab)
stdres(xtab)
```

No patterns stand out in particular.


