---
title: "Multiple correspondence analysis of BSL questions"
author: "Bodo Winter"
date: "04/02/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup

Load packages:

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(effsize) # for Cohen's d
library(FactoMineR) # for Multiple Correspondence Analysis
library(factoextra) # for interpreting MCA
```

Package versions:

```{r}
packageVersion('FactoMineR')
packageVersion('factoextra')
```

Load helper functions:

```{r}
source('question_helper_functions.R')
```

Load cleaned data, which is the output from the `preprocessing.Rmd` script:

```{r, message = FALSE, warning = FALSE}
df <- read_csv('../data/BSL_questions_cleaned.csv')
```


## Exclusions

Following the "turning points" in the `descriptive_statistics.Rmd` file we need to exclude constructed actions as well as "other" and "alternative" questions:

```{r}
df <- filter(df,
             has_CA == 'no CA',
             semantic %in% c('Content', 'Polar'))
```

## Multiple correspondence analysis

Create a vector that contains all the form variables.

```{r}
pred_vector <- c('any_head_movement', 'any_head_nod', 'any_head_forward',
                 'any_head_backward', 'any_head_sideward', 'any_head_shake',
                 'any_head_up', 'any_head_down', 'any_eye_brow',
                 'any_eye_brow_up', 'any_eye_brow_down',
                 'any_eye_brow_squint', 'any_eye_brow_wide',
                 'mouthing_question',
                 'manual', 'nonmanual')
```

Create a data frame with all the active variables that will be used in the MCA:

```{r}
df_active <- df[, pred_vector]
```

Run the MCA:

```{r}
BSL_MCA <- MCA(df_active, graph = FALSE)
```

Make a scree plot of this:

```{r, fig.width = 8, fig.eight = 6}
scree_plot <- fviz_screeplot(BSL_MCA, addlabels = TRUE,
                             ylim = c(0, 20))

# Plot & save:

scree_plot
ggsave(plot = scree_plot, file = '../figures/MCA_scree_plot.pdf',
       width = 10, height = 6)
```

Correlation between variables and dimensions:

```{r, fig.width = 12, fig.height = 6}
MCA_vars <- fviz_mca_var(BSL_MCA, choice = 'mca.cor',
                         repel = TRUE,
                         ggtheme = theme_minimal())

# Plot & save:

MCA_vars
ggsave(plot = MCA_vars, file = '../figures/MCA_vars.pdf',
       width = 10, height = 6)
```

Visualize contributions of the different variables to the first three dimensions:

```{r, fig.width = 12, fig.height = 6}
dim1_contribution <- fviz_contrib(BSL_MCA, choice = 'var', axes = 1, top = 15)
dim2_contribution <- fviz_contrib(BSL_MCA, choice = 'var', axes = 2, top = 15)
dim3_contribution <- fviz_contrib(BSL_MCA, choice = 'var', axes = 3, top = 15)

arrange(dim1_contribution$data, desc(contrib))
arrange(dim2_contribution$data, desc(contrib))

# Plot & save:

dim1_contribution
dim2_contribution
dim3_contribution
ggsave(plot = dim1_contribution, file = '../figures/pdf/MCA_dim1_contributions.pdf',
       width = 10, height = 6)
ggsave(plot = dim2_contribution, file = '../figures/pdf/MCA_dim2_contributions.pdf',
       width = 10, height = 6)
ggsave(plot = dim3_contribution, file = '../figures/pdf/MCA_dim3_contributions.pdf',
       width = 10, height = 6)
ggsave(plot = dim1_contribution, file = '../figures/png/MCA_dim1_contributions.png',
       width = 10, height = 6)
ggsave(plot = dim2_contribution, file = '../figures/png/MCA_dim2_contributions.png',
       width = 10, height = 6)
ggsave(plot = dim3_contribution, file = '../figures/png/MCA_dim3_contributions.png',
       width = 10, height = 6)
```

The first one is clearly: mouthing + manuals, which go together....

Let's look at the precise values:

```{r}
BSL_MCA_var <- get_mca_var(BSL_MCA)

round(BSL_MCA_var$contrib, 2)
```

Double-check that these are actually percentages out of the definition of the dimension:

```{r}
colSums(BSL_MCA_var$contrib)
```

Yes they are, column-wise things add up to 100%.

Extract individuals:

```{r}
MCA_inds <- get_mca_ind(BSL_MCA)

# Extract coordinates:

inds_coords <- as_tibble(MCA_inds$coord) %>% 
  rename(Dim1 = `Dim 1`,
         Dim2 = `Dim 2`,
         Dim3 = `Dim 3`,
         Dim4 = `Dim 4`,
         Dim5 = `Dim 5`)

# Append semantic type labels:

inds_coords <- bind_cols(inds_coords,
                         select(df, semantic, question, social_action))

# Check:

inds_coords
```

Make a plot of the individuals. If you swap "Q_semantic" with "Question" or "Q_social_action" you can get graphs for that.

```{r}
# Plot basics:

coord_p <- inds_coords %>%
  ggplot(aes(x = Dim1, y = Dim2, fill = semantic))

# Geoms:

coord_p <- coord_p +
  geom_vline(aes(xintercept = 0), linetype = 2) +
  geom_hline(aes(yintercept = 0), linetype = 2) +
  geom_point(size = 3, alpha = 0.75, col = 'black', pch = 21)

# Scales and axes:

coord_p <- coord_p +
  xlab('Dimension 1') +
  ylab('Dimension 2') +
  scale_fill_viridis_d(guide = guide_legend(reverse = TRUE)) +
  theme_BSL() +
  theme(legend.position = 'bottom',
        legend.title = element_blank(),
        axis.text.y = element_text(face = 'bold'))

# Plot & save:

coord_p
ggsave(plot = coord_p, file = '../figures/pdf/MCA_2D_solution_tokens.pdf',
       width = 8, height = 6)
ggsave(plot = coord_p, file = '../figures/png/MCA_2D_solution_tokens.png',
       width = 8, height = 6)
```

Check the dimension description:

```{r}
MCA_description <- dimdesc(BSL_MCA, axes = c(1, 2))

# Check them:

MCA_description[[1]]
MCA_description[[2]]
```

Reinforces the view that the first dimension is for manual + mouthing and the second is eyes & eye brows plus head.

Check which ones of the five dimensions differentiate `semantic` (content/polar questions):

```{r}
inds_coords |> 
  group_by(semantic) |> 
  summarize(M_dim1 = mean(Dim1),
            M_dim2 = mean(Dim2),
            M_dim3 = mean(Dim3),
            M_dim4 = mean(Dim4),
            M_dim5 = mean(Dim5))
```

See the effect sizes for this:

```{r}
with(inds_coords, cohen.d(Dim1 ~ semantic))
with(inds_coords, cohen.d(Dim2 ~ semantic))
with(inds_coords, cohen.d(Dim3 ~ semantic))
with(inds_coords, cohen.d(Dim4 ~ semantic))
with(inds_coords, cohen.d(Dim5 ~ semantic))
```

Ok, so the other dimensions really don't differentiate between polar and content questions. It's only Dimension 1.



